---
title: "Pathways to college"
subtitle: "Temuco Conference"
author: "Alberto F Cabrera"
date: 2023-10-16
format: 
  html:
    code-link: true
    code-fold: true
    self-contained: true
    embed-resources: true
    toc: true
    theme: Zephyr
    number-sections: false
knirt:
  opts_chunk: 
    warning: false
    editor_options: 
    chunk_output_type: inline
editor: visual
---

## Loading libraries

```{r}
#| label: loading_libraries
#| warning: false
#| echo: false

library(tidyverse)
library(gt)
library(gtExtras)
library(gtsummary)
library(gtools)
library(dplyr)
library(munsell)
library(ggthemes)



```

National Center for Education Statistics. Digest of Education Statistic's [table 303.10](https://nces.ed.gov/programs/digest/d22/tables/dt22_303.10.asp) Total fall enrollment in degree-granting post-secondary institutions, by attendance status, sex of student, and control of institution: Selected years, 1947 through 2031 (Estimated).

```{r}
#| label: retrieving_table_303_1
#| warning: false
#| echo: true

library(tidyverse)
library(haven)
library(janitor)

Enrollment_PSE_1947_2023 <- read_dta("Enrollment_PSE_1947-2023.dta") %>% 
  clean_names()

```

Enrollment in postsecondary education 1947-2023

```{r}
#| label: postsecondary_enrollment
#| warning: false
#| echo: true

Enrollment_PSE_1947_2023  %>% 
  filter(as.numeric(enr_th) <= "2010") %>% 
  ggplot(aes(x = year, y = enr_th)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(1973, 2023, 5)) +
  scale_y_continuous(breaks = seq(10000, 21000, 1000)) +
  labs(title = "Matrícula en educación superior (en miles)",
       subtitle = "1973 - 2023",
       caption = "Basado en la tabla 303.10 Compendio de Estadísticas Educativas 16 de octubre de 2023",
       y = element_blank(), 
       x = element_blank() ) +
  geom_text(x = 2017,
            y = 18500,
            label = "COVID-19",
            color = "red",
            size = 3.1,
            show.legend = FALSE) +
  geom_vline(xintercept = 2020, linetype = "dashed",
             color = "red", size = 1) +
  theme(
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.subtitle = element_text(size = 15, hjust = 0.5),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank()) +
  theme_classic()
  


```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
