---
title: "transfer_self_efficacy"
author: "Alberto F Cabrera"
date: 10-23-2024
format:
  html:
    code-link: true
    code-fold: true
    self-contained: true
    embed-resources: true
    toc: true
    theme: Zephyr
    number-sections: false
execute: 
  warning: false
  message: false
  echo: false
editor_options: 
  chunk_output_type: console
---

# Transfer Self-Efficacy

## Digest of Education Statistics

Relying on [Digest of Education Statistics](https://nces.ed.gov/programs/digest/current_tables.asp) to estimate high school graduates enrollment trends.

```{r loading_libraries}
#|echo: true
#|warning: false

library(tidyverse)
library(gt)
library(gtExtras)
library(gtsummary)
library(gtools)
library(gfonts)
library(janitor)
library(Hmisc)
library(expss)

```

## Percent of High School graduates enrolled in postsecondary education

Table 302.10. Number of recent high school completers and percent enrolled in college, by sex and level of institution: 1960 through 2022.

This table was retrieved on October 10, 2024. The table was prepared on December 2023.

### Retrieving and preparing database

Retrieving the excel file table302_2 HS enrollment in percentages. Renaming the variables and creating a pivot_longer excel file.

```{r hs_graduates_pse}
#| echo: true
#| warning: false

library(readxl)
table302_1_hs_enrollment <- read_excel("~/Documents/2024 Economic Inquiry in Education - EDU-633-003/data/tabn302.1_hs_enrollment.xlsx") |> 
  as_tibble()


table302_1_hs_enrollment <-  table302_1_hs_enrollment |> 
  rename(All = per_tot) |> 
  rename('Two year' = per_2yr) |> 
  rename('Four year' = per_4yr)
  

table302_1_hs_enrollment <- table302_1_hs_enrollment |> 
  pivot_longer(cols = 2:4, 
               names_to = "institution",
               values_to = "percent") |> 
  mutate(year = as.numeric(year),
         percent = round(percent,digits = 1)) |> 
  filter(year >= 2000)


```

Plotting percentage of high school graduates enrolling in postsecondary education.

```{r plot_hs_grads_enrolling}
#|echo: true
#|warning: false

#Setting up colors & assigning a unique color value to type of PSE

#thematic::okabe_ito(3)

colors_ito <- c('grey9', '#009E73', "#0072B2")

names(colors_ito) <- unique(table302_1_hs_enrollment$institution)

title <- glue::glue(
  'Percentage of HS graduates enrolling in ',
  '<span style="color:{colors_ito["Four year"]};">4-year</span> ',
  'and ',
  '<span style="color:{colors_ito["Two year"]};">2-year</span> ',
  'colleges')

subtitle <-  "(2000 - 2022)"
caption <- "Based on table 302.10 Digest of Education Statistics"


plot <- table302_1_hs_enrollment |> 
  ggplot(aes(x = year, y = percent, color = institution)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 2020, linetype = "dashed",
             color = '#E69F00', linewidth = 0.5)  +
  scale_x_continuous(breaks = seq(2000, 2023, 3)) +
  labs(title = title,
       subtitle = subtitle,
       caption = caption,
       x = element_blank(),
       y = element_blank(),
       color = " ") +
  geom_text(x = 2018.7,
            y = 55,
            label = "COVID-19",
            color = '#E69F00',
            size = 3.1) +
  geom_text(x = 2011,
            y = 71,
            label = "Overall",
            color = "grey9",
            size = 3.5) +
  theme_minimal(
    base_size = 12,
    base_family = "Source Sans Pro") +
  scale_color_manual(values = colors_ito) +
  theme(
    legend.position = 'none',
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = 'grey50'),
    plot.title = ggtext::element_markdown(
      family = 'Merriweather',
      size = rel(1.2),
      margin = margin(b = 0.3, unit = 'cm')
    ),
    plot.subtitle = ggtext::element_markdown(
      family = 'Merriweather',
      size = rel(1.1),
      margin = margin(b = 0.3, unit = 'cm')
    ),
    plot.caption = element_text(size = rel(0.8)) )


# Convert ggplot to interactive plot with plotly
plotly::ggplotly(plot, width = 800, height = 600)


```
