---
title: "transfer_self_efficacy"
author: "Alberto F Cabrera"
date: 10-23-2024
format:
  html:
    code-link: true
    code-fold: true
    self-contained: true
    embed-resources: true
    toc: true
    theme: Zephyr
    number-sections: false
execute: 
  warning: false
  message: false
  echo: false
editor_options: 
  chunk_output_type: console
---

# Transfer Self-Efficacy

## Digest of Education Statistics

Relying on [Digest of Education Statistics](https://nces.ed.gov/programs/digest/current_tables.asp) to estimate high school graduates enrollment trends.

```{r loading_libraries}
#|echo: true
#|warning: false

library(tidyverse)
library(gt)
library(gtExtras)
library(gtsummary)
library(gtools)
library(gfonts)
library(janitor)
library(Hmisc)
library(expss)
library(ggiraph)

```

## Percent of High School graduates enrolled in postsecondary education

Table 302.10. Number of recent high school completers and percent enrolled in college, by sex and level of institution: 1960 through 2022.

This table was retrieved on October 10, 2024. The table was prepared on December 2023.

### Retrieving and preparing database

Retrieving the excel file table302_2 HS enrollment in percentages. Renaming the variables and creating a pivot_longer excel file.

```{r hs_graduates_pse}
#| echo: true
#| warning: false

library(readxl)
table302_1_hs_enrollment <- read_excel("data_raw/tabn302.1_hs_enrollment.xlsx") |> 
  as_tibble()


# library(readxl)
# table302_1_hs_enrollment <- read_excel("~/Documents/2024 Economic Inquiry in Education - EDU-633-003/data/tabn302.1_hs_enrollment.xlsx") |> 
#   as_tibble()
# 

table302_1_hs_enrollment <-  table302_1_hs_enrollment |> 
  rename(All = per_tot) |> 
  rename('Two year' = per_2yr) |> 
  rename('Four year' = per_4yr)
  

table302_1_hs_enrollment <- table302_1_hs_enrollment |> 
  pivot_longer(cols = 2:4, 
               names_to = "institution",
               values_to = "percent") |> 
  mutate(year = as.numeric(year),
         percent = round(percent,digits = 1)) |> 
  filter(year >= 2000)


```

Plotting percentage of high school graduates enrolling in postsecondary education.

```{r plot_hs_grads_enrolling}
#|echo: true
#|warning: false

#Setting up colors & assigning a unique color value to type of PSE

#thematic::okabe_ito(3)

colors_ito <- c('grey9', '#009E73', "#0072B2")

names(colors_ito) <- unique(table302_1_hs_enrollment$institution)

title <- glue::glue(
  'Percentage of HS graduates enrolling in ',
  '<span style="color:{colors_ito["Four year"]};">4-year</span> ',
  'and ',
  '<span style="color:{colors_ito["Two year"]};">2-year</span> ',
  'colleges')

subtitle <-  "(2000 - 2022)"
caption <- "Based on table 302.10 Digest of Education Statistics"


plot <- table302_1_hs_enrollment |> 
  ggplot(aes(x = year, y = percent, color = institution)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 2020, linetype = "dashed",
             color = '#E69F00', linewidth = 0.5)  +
  scale_x_continuous(breaks = seq(2000, 2023, 3)) +
  labs(title = title,
       subtitle = subtitle,
       caption = caption,
       x = element_blank(),
       y = element_blank(),
       color = " ") +
  geom_text(x = 2018.7,
            y = 55,
            label = "COVID-19",
            color = '#E69F00',
            size = 3.1) +
  geom_text(x = 2011,
            y = 71,
            label = "Overall",
            color = "grey9",
            size = 3.5) +
  theme_minimal(
    base_size = 12,
    base_family = "Source Sans Pro") +
  scale_color_manual(values = colors_ito) +
  theme(
    legend.position = 'none',
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = 'grey50'),
    plot.title = ggtext::element_markdown(
      family = 'Merriweather',
      size = rel(1.2),
      margin = margin(b = 0.3, unit = 'cm')
    ),
    plot.subtitle = ggtext::element_markdown(
      family = 'Merriweather',
      size = rel(1.1),
      margin = margin(b = 0.3, unit = 'cm')
    ),
    plot.caption = element_text(size = rel(0.8)) )


# Convert ggplot to interactive plot with plotly
plotly::ggplotly(plot, width = 900, height = 500)


```

## Interactive figure using `ggiraph()`

This version adapts Albert Rapp's example.

In this modified script:

`geom_point_interactive()` and `geom_line_interactive()` from ggiraph allow adding tooltips for interactivity.

Adjust `width_svg`and `height_svg` parameters in girafe() to fit your preferred display dimensions.

This setup makes the plot interactive, allowing users to hover over points or lines to see additional information. Adjust the data and titles (subtitle and caption) as needed.

```{r ggiraph_version_albert_rapp}
#| echo: false
#| warning: false

library(tidyverse)
library(ggiraph)

#Setting up colors & assigning a unique color value to type of PSE

#thematic::okabe_ito(3)

colors_ito <- c('grey9', '#009E73', "#0072B2")

names(colors_ito) <- unique(table302_1_hs_enrollment$institution)


title <- glue::glue(
  'Percentage of HS graduates enrolling in ',
  '<span style="color:{colors_ito["Four year"]};">4-year</span> ',
  'and ',
  '<span style="color:{colors_ito["Two year"]};">2-year</span> ',
  'colleges')

subtitle <-  "(2000 - 2022)"
caption <- "Based on table 302.10 Digest of Education Statistics"


# Create the ggplot with ggiraph interactive elements

plot <- table302_1_hs_enrollment |> 
  mutate(tooltip_label = glue::glue("Year: {year}<br>Enrollment: {percent}%")) |> 
  ggplot(aes(x = year, y = percent, color = institution,
             linetype = institution)) +
  geom_point_interactive(
    aes(tooltip = tooltip_label), size = 3.5) +
  geom_line_interactive(
    aes(tooltip = paste0("Institution: ", institution)), size = 1) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "#E69F00", linewidth = 0.5) +
  scale_x_continuous(breaks = seq(2000, 2023, 2)) +
  labs(
    title = title,
    subtitle = subtitle,
    caption = caption,
    x = NULL,
    y = NULL,
    color = " "
  ) +
  geom_text(x = 2018.3, y = 55, label = "COVID-19", color = "#E69F00", size = 2.5) +
  geom_text(x = 2012.5, y = 71, label = "Overall", color = "grey9", size = 2.5) +
  theme_minimal(base_size = 14, base_family = "Source Sans Pro") +
  scale_color_manual(values = colors_ito) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "grey50"),
    plot.title = ggtext::element_markdown(
      family = "Merriweather", size = rel(1.1), margin = margin(b = 0.2, unit = "cm")),
    plot.subtitle = ggtext::element_markdown(
      family = "Merriweather", size = rel(.8), margin = margin(b = 0.2, unit = "cm")),
    plot.caption = element_text(size = rel(0.5))
  )

#Changing the border

girafe(
  ggobj = plot,
  width_svg = 8,
  height_svg = 5,
  options = list(
    opts_hover(css = ''),
    opts_hover_inv(
      css = girafe_css(
        css = 'opacity:0.4;',
        line = 'opacity:0.075;'
      )
    ),
    opts_tooltip(
      css = htmltools::css(
        background = 'white',
        border = '2px solid black;',
        padding = '10px',
        font_weight = 600,
        font_size = '14p'))))




```

```{r ggiraph_version}
#| echo: false
#| warning: false

library(ggiraph)

title <- glue::glue(
  'Percentage of HS graduates enrolling in ',
  '<span style="color:{colors_ito["Four year"]};">4-year</span> ',
  'and ',
  '<span style="color:{colors_ito["Two year"]};">2-year</span> ',
  'colleges')

subtitle <-  "(2000 - 2022)"
caption <- "Based on table 302.10 Digest of Education Statistics"


# Create the ggplot with ggiraph interactive elements

plot <- table302_1_hs_enrollment |> 
  ggplot(aes(x = year, y = percent, color = institution)) +
  geom_point_interactive(
    aes(tooltip = paste0("Year: ", year, "<br>Enrollment: ", percent, "%")), size = 3) +
  geom_line_interactive(
    aes(tooltip = paste0("Institution: ", institution)), size = 1) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "#E69F00", linewidth = 0.5) +
  scale_x_continuous(breaks = seq(2000, 2023, 3)) +
  labs(
    title = title,
    subtitle = subtitle,
    caption = caption,
    x = NULL,
    y = NULL,
    color = " "
  ) +
  geom_text(x = 2018.3, y = 55, label = "COVID-19", color = "#E69F00", size = 2.5) +
  geom_text(x = 2012.5, y = 71, label = "Overall", color = "grey9", size = 2.5) +
  theme_minimal(base_size = 10, base_family = "Source Sans Pro") +
  scale_color_manual(values = colors_ito) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "grey50"),
    plot.title = ggtext::element_markdown(
      family = "Merriweather", size = rel(1), margin = margin(b = 0.2, unit = "cm")),
    plot.subtitle = ggtext::element_markdown(
      family = "Merriweather", size = rel(.8), margin = margin(b = 0.2, unit = "cm")),
    plot.caption = element_text(size = rel(0.8))
  )

# Convert ggplot to ggiraph plot
girafe(ggobj = plot, width_svg = 5, height_svg = 3)


```

Additional changes to the chart by making the lines interactive as well and lessening the colors of the other continents when highlighting one of them.

```{r ggiraph_interactive_last_option}
#| echo: false
#| warning: false

library(ggiraf)

#Setting up colors & assigning a unique color value to type of PSE

#thematic::okabe_ito(3)

colors_ito <- c('grey9', '#009E73', "#0072B2")

names(colors_ito) <- unique(table302_1_hs_enrollment$institution)

title <- glue::glue(
  'Percentage of HS graduates enrolling in ',
  '<span style="color:{colors_ito["Four year"]};">4-year</span> ',
  'and ',
  '<span style="color:{colors_ito["Two year"]};">2-year</span> ',
  'colleges')

subtitle <-  "(2000 - 2022)"
caption <- "Based on table 302.10 Digest of Education Statistics"


# Create the ggplot with ggiraph interactive elements

plot <- table302_1_hs_enrollment |> 
  ggplot(aes(x = year, y = percent, color = institution)) +
  geom_point_interactive(
    aes(tooltip = paste0("Institution: ",institution, "<br>Year: ", year, "<br>Enrollment: ", percent, "%"),
        data_id = institution), size = 3) +
  geom_line_interactive(
    aes(#tooltip = paste0("Institution: ", institution),
        data_id = institution), size = 1) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "#E69F00", linewidth = 0.5) +
  scale_x_continuous(breaks = seq(2000, 2023, 2)) +
  labs(
    title = title,
    subtitle = subtitle,
    caption = caption,
    x = NULL,
    y = NULL,
    color = " "
  ) +
  geom_text(x = 2018.3, y = 55, label = "COVID-19", color = "#E69F00", size = 2.5) +
  geom_text(x = 2012.5, y = 71, label = "Overall", color = "grey9", size = 2.5) +
  theme_minimal(base_size = 10, base_family = "Source Sans Pro") +
  scale_color_manual(values = colors_ito) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "grey50"),
    plot.title = ggtext::element_markdown(
      family = "Merriweather", size = rel(1), margin = margin(b = 0.2, unit = "cm")),
    plot.subtitle = ggtext::element_markdown(
      family = "Merriweather", size = rel(.8), margin = margin(b = 0.2, unit = "cm")),
    plot.caption = element_text(size = rel(0.8))
  )

# Convert ggplot to ggiraph plot
girafe(ggobj = plot, 
       width_svg = 5, 
       height_svg = 3,
       options = list(
         opts_hover(css = ''),
         opts_hover_inv(
           css = girafe_css(
             css = 'opacity:0.4;',
             line = 'opacity:0.075;' )),
         opts_hover_inv(css = 'opacity:0.2;'),
         opts_tooltip(
           css = htmltools::css(
             background = 'white',
             border = '2px solid black;',
             padding = '10px',
             font_weight = 600,
             font_size = '14p'
           )
         )
       ))


```

A new version with a longer legend

```{r girafe_longer_legend}
#| echo: false
#| warning: false


library(tidyverse)
library(gt)
library(gtExtras)
library(gtsummary)
library(gtools)
library(gfonts)
library(janitor)
library(Hmisc)
library(expss)
library(ggiraph)
library(readxl)

table302_1_hs_enrollment <- read_excel("~/Documents/2024 Economic Inquiry in Education - EDU-633-003/data_raw/tabn302.1_hs_enrollment.xlsx") |> 
  as_tibble()

table302_1_hs_enrollment <-  table302_1_hs_enrollment |> 
  rename('all' = per_tot) |> 
  rename('2-year' = per_2yr) |> 
  rename('4-year' = per_4yr)
  

table302_1_hs_enrollment <- table302_1_hs_enrollment |> 
  pivot_longer(cols = 2:4, 
               names_to = "institution",
               values_to = "percent") |> 
  mutate(year = as.numeric(year),
         percent = round(percent,digits = 1)) |> 
  filter(year >= 2000)


#Setting up colors & assigning a unique color value to type of PSE

#thematic::okabe_ito(3)

colors_ito <- c('grey9', '#009E73', "#0072B2")

names(colors_ito) <- unique(table302_1_hs_enrollment$institution)
title <- glue::glue(
  'Percentage of HS graduates enrolling in ',
  '<span style="color:{colors_ito["4-year"]};">**4-year**</span> ',
  'and ',
  '<span style="color:{colors_ito["2-year"]};">**2-year**</span> ',
  'colleges')

subtitle <-  "(2000 - 2022)"
caption <- "Based on table 302.10 *Digest of Education Statistics*"

annotations <- data.frame(
  x = c(2018.3, 2012.5),
  y = c(55, 71),
  label = c("COVID-19", "Overall"),
  color = c("#E69F00", "grey9")
)


# Create the ggplot with ggiraph interactive elements

plot <- table302_1_hs_enrollment |> 
  mutate(tooltip_label = glue::glue("{percent}% of HS grads enrolled<br> in {institution} colleges in {year}")) |> 
  ggplot(aes(x = year, y = percent, color = institution,
             linetype = institution)) +
  geom_point_interactive(
    aes(tooltip = tooltip_label, data_id = institution), size = 3.5) +
  geom_line_interactive(
    aes(data_id = institution), size = 1) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "#E69F00", linewidth = 0.5) +
  scale_x_continuous(breaks = seq(2000, 2023, 2)) +
  labs(
    title = title,
    subtitle = subtitle,
    caption = caption,
    x = NULL,
    y = NULL,
    color = " "
  ) +
  geom_text(data = annotations, aes(x = x, y = y, label = label, color = color), size = 3.5) +
  theme_minimal(base_size = 14, base_family = "Source Sans Pro") +
  scale_color_manual(values = colors_ito) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "grey50"),
    plot.title = ggtext::element_markdown(
      family = "Merriweather", size = rel(1.1), margin = margin(b = 0.2, unit = "cm")),
    plot.subtitle = ggtext::element_markdown(
      family = "Merriweather", size = rel(.8), margin = margin(b = 0.2, unit = "cm")),
    plot.caption = ggtext::element_markdown(size = rel(0.5))
  )

#Changing the border

girafe(ggobj = plot, 
       width_svg = 8, 
       height_svg = 5,
       options = list(
         opts_hover(css = ''),
         opts_hover_inv(
           css = girafe_css(
             css = 'opacity:0.4;',
             line = 'opacity:0.075;' )),
         opts_hover_inv(css = 'opacity:0.2;'),
         opts_tooltip(
           css = htmltools::css(
             background = 'white',
             border = '2px solid black;',
             padding = '10px',
             font_weight = 600,
             font_size = '14p'
           )
         )
       ))





```
